

This thesis has been an enriching and tough journey during it I was surrounded by many supportive people, to whom I would like to express my gratitude. 
First and foremost, I am indebted to Prof.~Dr.~Melanie Herschel who advised, supported me during those four years. 
Her constant support and guidance have been prominent throughout this journey, and I am extremely grateful for that.


%I would like also to express my sincere gratitude to Prof.~Dr.~Carsten Binnig for accepting to evaluate my research work. 
%%Professor xxx and Professor xxx for being my thesis reviewers and for their thoughtful and encouraging comments. Professor xxxx and Professor xxxx for being my thesis examiners and for their relevant remarks.


I am thankful to my former colleagues from the Data Engineering group at the University of Stuttgart for their valuable support, technical discussions, and making the time here pleasant.
I owe my warmest affection to all the members of the computer science department of University of Stuttgart. Especially to Eva for her kind help and assistance. 

%{\color{Fuchsia}
%I want also to to thank the German Research Foundation (DFG) for funding my research within the project ?Quantitative methods for visual computing? (SFB-TRR 161). I express my gratitude to all the project members for their collaboration, feedback, and discussions within interdisciplinary projects that allowed us to apply and evaluate our research.
I want also to thank the German Research Foundation (DFG) for funding my research within the project ``Quantitative methods for visual computing'' (SFB-TRR 161). I express my gratitude to all the project members for their feedback and discussions made during our collaboration.
%}

Finally, I would like to acknowledge my family for their support. I thank my mother Halima and my father Ali, who were always there to provide unconditional support.  I also thank my sister Imen and my brother Karim for their support and encouragement.