\section{Introduction}
%As motivated in the previous chapters, we propose in this thesis a novel framework for
This chapter formalizes our framework for holistic recommendation-based visual data exploration. Furthermore, we present the set of features and functionality that our framework makes available to the users to facilitate their exploration tasks. 



We introduce our framework in Section~\ref{sec:frame-overview}. 
In Section~\ref{sec:frame-comp}, we describe the main modules and components present in this framework.
Subsequently, we describe in Section~\ref{sec:frame-navi} the set of possible navigations that can be made using our framework when exploring data visually.
Furthermore, we discuss in Section~\ref{sec:frame-process} the process of visual data exploration that users can perform using an instance of our framework. 
%Finally, we conclude this chapter before introducing for the next chapter.
Note that this chapter relies partially on descriptions and exploration scenarios presented in our paper~\cite{BHBK18:Demo}.


%\section{Provenance-based visual data exploration framework}
\section{Framework overview}
\label{sec:frame-overview}
\input{chapterOverview/frameworkOverview}
%\vspace{-3em}
\section{Framework components}
\label{sec:frame-comp}
\input{chapterOverview/frameworkComponents}

\section{Framework navigations}
\label{sec:frame-navi}
\input{chapterOverview/frameworkNavigations}

\section{Framework implementation}
\label{sec:frame-process}
\input{chapterOverview/frameworkInstance}



\section{Conclusion}
This chapter introduced our holistic recommendation-based visual data exploration framework \framework{}. Besides that, we presented \prototype{}, an instance of our framework meant to explore data warehouses visually. 
In the next chapters, we continue describing our framework. Hence, we describe thoroughly two main modules present in our framework that are the \emph{provenance engine module} and the \emph{Recommendation engine module}. 
%\mel{Worth saying that while provenance module is general, recommendations are tailored to particular framework instance (and in our case EVLIN for DW). Briefly justify this, if possible.}
While the description of the \emph{provenance engine module} is general, the description of the \emph{Recommendation engine module} is tailored to a particular framework instance (in our case \prototype{} for visual exploration of data warehouse). The rationale behind that is to provide evidences about the feasibility of our proposed recommendations methods.







