 \section{Conclusion}
 

{\color{Fuchsia}In this chapter, we presented our scientific contributions underlying our visual data exploration framework \framework{}.
In particular, we discussed our provenance-based recommendation techniques including the content-based query recommendation, the collaborative-filtering query recommendation and the visualization recommendation.
Furthermore, we described also our approach proposed to quantify the interestingness of generated recommendations.

We performed also a set of evaluations to validate our contributions.  
 More specifically, we evaluated first our content-based query recommendation approach, our visualization recommendation solution and our methods quantifying recommended queries.  


Overall, our contributions were evaluated  both quantitatively using performance measurements and qualitatively with a user study on both synthetic and real data. 


Quantitative experiments show the feasibility and the efficiency of our proposed solutions for visual and interactive data exploration.
The qualitative evaluations was insightful as it shows a general satisfaction among users about query and visualization recommendations supported by our framework. 
It shows also that users find sometimes difficulties to distinguish between the interestingness of recommendations proposed using our framework.

Accordingly, we suggest in the next chapter further contributions to reinforce our framework {\color{Fuchsia}\framework{}}. Our extensions aim at improving the accuracy of recommendations' interestingness score to offer users possibility to perform a more efficient visual data exploration. }




