Parts of this thesis have been published in:

The remainder of this thesis manuscript is organized as follows:
{\color{Fuchsia}
\begin{itemize}
\item \textbf{H. Ben Lahmar}, M. Herschel, \emph{Provenance-based recommendations for visual data exploration}, in the USENIX Workshop on the Theory and Practice of Provenance, TaPP, 2017.
\item M. Herschel, R. Diestelkaemper, \textbf{H. Ben Lahmar}, \emph{A survey on provenance: what for? what form? what from?}, in the Very Large Databases Journal, VLDBJ, vol. 26, no. 6, pp. 881--906, 2017. 
\item \textbf{H. Ben Lahmar}, M. Herschel, M. Blumenschein, D.A  Keim, \emph{Provenance-based visual data exploration with EVLIN}, in the International Conference on Extending Database Technology, EDBT, pp. 686--689, 2018. 
\item C. Schulz, A. Zeyfang, M. Garderen, \textbf{H. Ben Lahmar}, M. Herschel, D. Weiskopf, \emph{Simultaneous visual analysis of multiple software hierarchies}, in the IEEE Working Conference on Software Visualization, VISSOFT, pp. 87--95, 2018. 
\item \textbf{H. Ben Lahmar} and M. Herschel, \emph{Structural summaries for visual provenance analysis}, in the workshop on Theory and Practice of Provenance, TaPP, 2019.
\item \textbf{H. Ben Lahmar} and M. Herschel, \emph{Towards integrating collaborative filtering in visual data exploration systems}, in the European Conference on Advances in Databases and Information Systems, ADBIS, 2019.

\item \textbf{H. Ben Lahmar} and M. Herschel, \emph{Collaborative filtering over evolution provenance data for interactive visual data exploration},Information System Journal, 2019 (under submission).
\item  V. Bruder, \textbf{H. Ben Lahmar}, M. Hlawatsch, S. Frey, M. Burch, D. Weiskopf, M. Herschel , T. Ertl, \emph{Volume-Based Large Dynamic Graph Analysis Supporting Evolution Provenance}, in the Multimedia Tools and Applications Journal, MTAP, 2019.\\
V. Bruder  took the major responsibilities of this paper and implemented a system for visual analysis of dynamic large graphs. I and M. Herschel contributed to the extension of this system. Indeed, we disseminate our evolution provenance model in this system providing users thereby with more features such as recall explored parts and documentation of exploration process.


\end{itemize}}